Practical Optional Types for Clojure

### 2015-02-19 small.tex
use minted for syntax highlighting
Abstract.

\documentclass[preprint,9pt,twocolumn]{sigplanconf}

@@

\usepackage{common}
\include{bibliography.bib}

@@

\begin{abstract}
Clojure is a dynamically typed language hosted on the Java 
Virtual Machine.
Typed Racket is a valuable starting point for
a gradual type system that targets Clojure.
Building a similar type system for a new language gives the
designer some flexibility to repurpose and extend features.
This paper gives an overview of Typed Clojure, concentrating
on the extensions and differences from Typed Racket. We also
show where Typed Racket's features were particularly useful
for type checking non-trivial Clojure idioms.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\keywords
optional types, Clojure

>>>>

\documentclass[preprint,9pt,twocolumn,nocopyrightspace,authoryear]{sigplanconf}

@@

\usepackage{common}
\usepackage{minted}
\include{bibliography.bib}

@@

\begin{abstract}
Typed Clojure is an optional type system for Clojure, a dynamic
language in the Lisp family that targets the JVM. Typed Clojure's type
system build on the design of Typed Racket, repurposing in particular
\emph{occurence typing}, its approach to statically reasoning about
predicate tests. However, in adapting the type system to Clojure,
changes and extensions are required to accomodate both additional
language features and  idioms used by Clojure programmers.

In this paper, we describe Typed Clojure and present these type system
extensions, focusing on three features widely used in Clojure. First, Java
interoperability is central to Clojure's mission but introduces
challenges such as ubiquitous \texttt{null}; Typed Clojure handles
Java interoperability while ensuring the absence of null-pointer
exceptions in typed programs. Second, Clojure programmers
idiomatically use immutable dictionaries for data structures; Typed
Clojure handles this in the type system with multiple forms of
heterogenous dictionary types. Third, multimethods provide extensible
operations, and their Clojure semantics turns out to have a surprising
synergy with the underlying occurence typing framework.

We provide a formal model of the Typed Clojure type system
incorporating these and other features, with a proof of
soundness. Additionally, Typed Clojure is now in use by numerous
corporations and developers working with Clojure, and we report on 
experience with the system and its lessons for the future.

%% Clojure is a dynamically typed language hosted on the Java 
%% Virtual Machine.
%% Typed Racket is a valuable starting point for
%% a gradual type system that targets Clojure.
%% Building a similar type system for a new language gives the
%% designer some flexibility to repurpose and extend features.
%% This paper gives an overview of Typed Clojure, concentrating
%% on the extensions and differences from Typed Racket. We also
%% show where Typed Racket's features were particularly useful
%% for type checking non-trivial Clojure idioms.
\end{abstract}

%% \category{CR-number}{subcategory}{third-level}

%% \keywords
%% optional types, Clojure
